Stochastic Convolutional Sparse Coding

### 2018-11-15 cvpr2019/SCSCmodel.tex
slight reshuffle

\section{Stochastic Convolutional Sparse Coding}
\subsection{The Model}
We first define $\code = [\code_1, \dots, \code_K] \in \mathbb{R}^{DK}$.
Based on the strong sparsity of the codes $\code$, we propose to
implement the CSC learning problem~\eqref{eq:CSCmodel} iteratively,
where at $\ite$-th iteration we only consider a random subset of the codes, denoted as $\subcode^t$.
We subsample the sparse codes $\code$ following a Bernoulli distribution with
probability $p$. Herein, $\subcode^t \in \mathbb{R}^{pDK}$, which satisfies:
\begin{equation}
    \subcode^t = \mask^{\ite} \code^\ite,
\end{equation}
where $\mask^{\ite}$ is a $\mathbb{R}^{pDk \times DK}$ random binary matrix redrawn at current iteration, with $1$s on the entries corresponding to the sampled codes and $0$s elsewhere. Concretely, it projects the codes $\code^\ite$ to a random subspace, retaining the sampled ones and filtering out the others. In the case of $p=1$, the proposed model is identical with the classical CSC model, and $\mask$ is simply an identity matrix. When $p<1$, the algorithm only solves a subset of the codes at chosen positions in each iteration, and accordingly, the update of dictionary $\filter$ is based on the selected
portion of the codes $\subcode$. Similar to solving the classical CSC problem, we can apply a coordinate descent algorithm, alternating on
subproblems of $\subcode$ and $\filter$, to tackle the bi-convex optimization problem. Specifically, the modified
minimization problem for $\subcode$ can be formulated as:
\begin{equation} \label{eq:updatingCode}
    \subcode^{\ite} = \minimize{\subcode} \frac{1}{2}\| \signal - \Filter^{\ite-1} (\mask^{\ite})^\top \subcode \|_2^2 + \lambda \| \subcode \|_1,
\end{equation}
where $\Filter^{\ite-1} = [\Filter_1^{\ite-1}, \dots, \Filter_K^{\ite-1}] \in \mathbb{R}^{D
  \times DK}$, which is constructed from the dictionary learned by
$(\ite-1)$-th iteration. The convolution operators are expressed as a matrix
multiplication so that $ \Filter \code = \sum_{k=1}^{K} \filter_k *
\code_k$. Therefore, each part of $\Filter$ is a Toeplitz
matrix. Due to the introduced subsampling matrix, the
convolution operator cannot be implemented in the Fourier
domain. However, owing to the subsampling
strategy, the number of variables that need to be computed for this
subproblem is $pDK$ instead of $DK$, which leads to a reduction
of the spatial-domain computation time by a factor of $\frac{1}{p}$.

After computing the subsampled codes $\subcode^\ite$, we can then project them
onto the original spatial support by
\begin{equation} \label{eq:upsampleCode}
    \code^\ite = (\mask^{\ite})^\top \subcode^\ite.
\end{equation}
Afterwards, the dictionary can be updated by:
\begin{equation} \label{eq:updatingFilter}
\begin{split}
   & \filter^{\ite} = \minimize{\filter} \frac{1}{2}\|\signal - \Code^{\ite} \filter \|_2^2 \\
   & \text{subject to}  ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\Code^{\ite}= [\Code_1^{\ite}, \dots, \Code_K^{\ite}]
\in \mathbb{R}^{D \times MK}$ is a concatenation of Toeplitz matrices,
and $\Code_k^{\ite}$ is constructed from the associated $\code_k^{\ite}$, $\filter=
[\filter_1,\dots,\filter_K] \in \mathbb{R}^{MK}$ such that $ \Code
\filter = \sum_{k=1}^{K} \filter_k * \code_k$. In typical CSC
settings, $M \ll D$, hence there is no need to perform subsampling on
the dictionary.

>>>>

\section{Stochastic Convolutional Sparse Coding}
\subsection{The Model}
We first define a stacked vector $\code = [\code_1, \dots, \code_K]
\in \mathbb{R}^{DK}$ for the codes, as well as a stacked matrix
$\Filter = [\Filter_1, \dots, \Filter_K] \in \mathbb{R}^{D \times DK}$
for the filter convolutions. The convolution operators are expressed
as a matrix multiplication so that $ \Filter \code = \sum_{k=1}^{K}
\filter_k * \code_k$.  Therefore, each part of $\Filter$ is a Toeplitz
matrix.

Based on the strong sparsity of the codes $\code$, we propose to
implement the CSC learning problem~\eqref{eq:CSCmodel} iteratively,
where at the $\ite$-th iteration we only consider a random subset of
the codes, denoted as $\subcode^t\in \mathbb{R}^{pDK}$.  We subsample
the sparse codes $\code$ following a Bernoulli distribution with
probability $p$. This subsampling process can be expressed as a matrix
operation
\begin{equation}
    \subcode^t = \mask^{\ite} \code^\ite,
\end{equation}
where $\mask^{\ite}$ is a $\mathbb{R}^{pDk \times DK}$ binary matrix,
where exactly one entry per row has a value of $1$, and the other
entries are $0$. For each iteration $\ite$, a different matrix
$\mask^{\ite}$ is generated randomly. This matrix projects the
codes $\code^\ite$ to a random subspace, retaining the sampled codes
and filtering out the others. In the case of $p=1$, the proposed model
is identical to the classical CSC model, and $\mask$ is simply an
identity matrix. When $p<1$, the algorithm only solves a subset of the
codes at chosen positions in each iteration, and accordingly, the
update of dictionary $\filter$ is based on the selected portion of the
codes $\subcode$. Similar to solving the classical CSC problem, we can
apply a coordinate descent algorithm, alternating on subproblems of
$\subcode$ and $\filter$, to tackle the bi-convex optimization
problem. Specifically, the modified minimization problem for
$\subcode$ can be formulated as:
\begin{equation} \label{eq:updatingCode}
    \subcode^{\ite} = \minimize{\subcode} \frac{1}{2}\| \signal - \Filter^{\ite-1} (\mask^{\ite})^\top \subcode \|_2^2 + \lambda \| \subcode \|_1,
\end{equation}
where $\Filter^{\ite-1}$ is composed of the dictionary learned in the
$(\ite-1)$-th iteration.
 Due to the introduced subsampling matrix, the convolution
operator cannot be implemented in the Fourier domain. However, owing
to the subsampling strategy, the number of variables that need to be
computed for this subproblem is $pDK$ instead of $DK$, which leads to
a reduction of the spatial-domain computation time by a factor of
$\frac{1}{p}$.

After computing the subsampled codes $\subcode^\ite$, we can then project them
onto the original spatial support by
\begin{equation} \label{eq:upsampleCode}
    \code^\ite = (\mask^{\ite})^\top \subcode^\ite.
\end{equation}
Afterwards, the dictionary can be updated by solving the optimization problem
\begin{equation} \label{eq:updatingFilter}
\begin{split}
   & \filter^{\ite} = \minimize{\filter} \frac{1}{2}\|\signal - \Code^{\ite} \filter \|_2^2 \\
   & \text{subject to}  ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\Code^{\ite}= [\Code_1^{\ite}, \dots, \Code_K^{\ite}]
\in \mathbb{R}^{D \times MK}$ is a concatenation of Toeplitz matrices,
and $\Code_k^{\ite}$ is constructed from the associated $\code_k^{\ite}$, $\filter=
[\filter_1,\dots,\filter_K] \in \mathbb{R}^{MK}$ such that $ \Code
\filter = \sum_{k=1}^{K} \filter_k * \code_k$. In typical CSC
settings, $M \ll D$, hence there is no need to perform subsampling on
the dictionary.
